---
title: "Homework 4"
author: Na Yun Cho
output: pdf_document
---


```{r}
library(ISLR)
library(mlbench)
library(caret)
library(rpart)
library(rpart.plot)
library(party)
library(partykit)
library(plotmo)
library(pROC)
library(randomForest)
library(ranger)
library(gbm)
library(pdp)
library(lasso2)
library(tidyverse) # data manipulation
library(ISLR) # data Problem 2
library(patchwork)
library(vip)
```

## 1(a)
Fit a regression tree with lpsa as the response and the other variables as predictors.
```{r, message=F}
set.seed(1)
data(Prostate)
Prostate <-na.omit(Prostate)

# partition the dataset 
trRows <- createDataPartition(Prostate$lpsa,
                              p =0.75, list =F)

tree1 <- rpart(formula = lpsa ~ . , 
               data = Prostate, subset = trRows,
               control = rpart.control(cp = 0))
printcp(tree1)
cpTable <- tree1$cptable
plotcp(tree1)

# tree using lowest cross validation error 
minErr <- which.min(cpTable[,4])
tree3 <- prune(tree1, cp = cpTable[minErr,1])
rpart.plot(tree3)

# tree using the 1SE rule 
tree4 <- prune(tree1, cp = cpTable[cpTable[,4]<cpTable[minErr,4]+cpTable[minErr,5],1][1])
rpart.plot(tree4)
```
Tree size of 6 corresponds to the lowest cross-validation error. Tree size of 4 is obtained using the 1 SE rule. 
Thus, the sizes are different. 

```{r}
# use caret to do cross validation
set.seed(1)
ctrl <- trainControl(method = "cv")
rpart.fit <- train(lpsa~., Prostate[trRows,], 
                   method = "rpart",
                   tuneGrid = data.frame(cp = exp(seq(-6,-2, length = 50))),
                   trControl = ctrl)
ggplot(rpart.fit, highlight = TRUE)
rpart.plot(rpart.fit$finalModel)
rpart.fit$bestTune
```
The optimal tree size is 4. 

## 1(b) plot of the final tree
```{r}
tree5 <- prune(tree1, cp = 0.05081357)
rpart.plot(tree5)
vip(tree5)
```
The final tree I chose has a size of 4. The interpretation of the bottom right terminal node is that for observations with `lcavol` (log cancer volume) that are greater than or equal to 2.5, the value of response (log prostate specific antigen) is predicted to be 4. 20% of the data correspond to this bottom right terminal node. 


## 1(c) bagging
```{r}
# perform bagging 
bagging <- ranger(lpsa~., Prostate[trRows,],
                  mtry = 8,
                  importance = "permutation",
                  min.node.size = 25,
                  scale.permutation.importance = TRUE)
barplot(sort(ranger::importance(bagging), decreasing = FALSE), 
        las = 2, horiz = TRUE, cex.names = 0.7,
        col = colorRampPalette(colors = c("darkred","white","darkblue"))(19))
```

For bagging, we do not have to tune the `mtry` parameter. 
The variable importance plot is shown above.
'lcavol' is the most important predictor, 'lweight' is the second most important one, and 'pgg45' is the third important predictor. Meanwhile, 'age' is the least important predictor. 

## 1(d) random forest 
```{r}
rf.grid <- expand.grid(mtry = 1:8, 
                       splitrule = "variance",
                       min.node.size = 1:30)
rf.fit <- train(lpsa~., Prostate[trRows,],
                method = "ranger",
                tuneGrid = rf.grid,
                trControl = ctrl)
ggplot(rf.fit, highlight = TRUE)


rf.final.per <- ranger(lpsa ~ . , 
                        Prostate[trRows,],
                        mtry = rf.fit$bestTune[[1]], 
                        splitrule = "variance",
                        min.node.size = rf.fit$bestTune[[3]],
                        importance = "permutation",
                        scale.permutation.importance = TRUE) 

barplot(sort(ranger::importance(rf.final.per), decreasing = FALSE), 
        las = 2, horiz = TRUE, cex.names = 0.7,
        col = colorRampPalette(colors = c("cyan","blue"))(19))
```
'lcavol' is the most important predictor, 'lweight' is the second most important predictor, and 'pgg45' is the third most important one. Meanwhile, 'age' is the least important predictor. 

## 1(e) boosting 
```{r}
set.seed(1)
gbm.grid <- expand.grid(n.trees = c(2000,3000,4000),
                        interaction.depth = 1:4,
                        shrinkage = c(0.001,0.003,0.005),
                        n.minobsinnode = c(1,10))
gbm.fit <- train(lpsa ~ . , 
                 Prostate[trRows,], 
                 method = "gbm",
                 tuneGrid = gbm.grid,
                 trControl = ctrl,
                 verbose = FALSE)

ggplot(gbm.fit, highlight = TRUE)
gbm.fit$bestTune

# variable importance
summary(gbm.fit$finalModel, las = 2, cBars = 19, cex.names = 0.6)
```
I can see that 'lcavol' is the most important variable. 'lweight' is the second important variable and 'pgg45' is the third important variable. 'gleason' is the least important variable. 

## 1(f) model selection
```{r}
resamp <- resamples(list(gbm = gbm.fit, rf = rf.fit, rpart = rpart.fit))
summary(resamp)

```
I would use ensemble methods rather than a single regression tree to predict PSA level, because ensemble methods show higher prediction accuracy. Among the ensemble methods, bagging is a special case of random forest in which mtry equals the total number of predictors, which is 8 in this assignment. According to the tuning process of random forest, the best mtry shows as 7, which is smaller than 8. This indicates random forest is better than bagging. Here, I compare regression tree, random forest, and boosting using cross validation. It shows that Boosting has lower mean cross-validation RMSE than that of random forest, which indicates that it has the best prediction accuracy. Therefore, I will choose the boosting model to predict PSA level.




